---
title: "geom_bar_delim"
author: "Evangeline Reynolds"
date: "4/24/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Puzzle: how to get delimited column of data to be split up, counted and returned as bar chart.

## Do it with base ggplot2

```{r}
library(tidyverse)

toy_survey <- data.frame(q1 = c('excitement;skepticism', 'skepticism;curiousity'))


toy_survey %>%
  pull(q1) %>%
  paste(collapse = ";") %>%
  str_split_1(";") %>%
  data.frame(x = .) %>%
  count(x) %>%
  mutate(y = n) %>%
  ggplot() +
  aes(x = x, y = y) +
  geom_col() + 
  labs(x = NULL,
       y = "Number of respondents indicating sentiment")

layer_data(last_plot())
```


# 2. try to extend

```{r}
compute_panel_bar_delim <- function(data,
                                     scales){

  data$responses %>%
    paste(collapse = ";") %>%
    str_split_1(";") %>%
    data.frame(cats = .) %>%
    count(cats) %>%
    rename(num_responding = n)

}


toy_survey %>%
  rename(responses = q1) %>%
  compute_panel_bar_delim()
```


```{r}
setup_data_function <- function(data, params){

  if(data$group[1] == -1){
    nrows <- nrow(data)
    data$group <- seq_len(nrows)
  }

  data

}



StatBardelim <- ggplot2::ggproto(
  `_class` = "StatBardelim",
  `_inherit` = ggplot2::Stat,
  # required_aes = c("responses"),
  compute_panel = compute_panel_bar_delim,
  default_aes = ggplot2::aes(x = after_stat(cats), 
                             y = after_stat(num_responding), 
                             group = after_stat(cats))#,
  # setup_data = setup_data_function
  # compute_goup = compute_group_bar_delim
)

geom_bar_delim <- function(
  mapping = NULL,
  data = NULL,
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatBardelim,  # proto object from Step 2
    geom = ggplot2::GeomCol,  # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

# try to use extension

```{r, error=T}
toy_survey %>%
  ggplot() +
  aes(responses = q1) +
  geom_bar_delim()
```



Backtrace:

  1. base (local) `<fn>`(x)

  2. ggplot2:::print.ggplot(x)

  4. ggplot2:::ggplot_build.ggplot(x)

  5. ggplot2:::by_layer(...)

 12. ggplot2 (local) f(l = layers[[i]], d = data[[i]])

# Some ideas

- maybe cat needs to be a factor?




# 2. try to extend

```{r}
compute_panel_bar_delim <- function(data,
                                     scales){

  data$responses %>%
    paste(collapse = ";") %>%
    str_split_1(";") %>%
    data.frame(cats = .) %>%
    count(cats) %>%
    mutate(cats = as.factor(cats)) %>% 
    rename(num_responding = n)

}


toy_survey %>%
  rename(responses = q1) %>%
  compute_panel_bar_delim()
```


```{r}
setup_data_function <- function(data, params){

  if(data$group[1] == -1){
    nrows <- nrow(data)
    data$group <- seq_len(nrows)
  }

  data

}



StatBardelim <- ggplot2::ggproto(
  `_class` = "StatBardelim",
  `_inherit` = ggplot2::Stat,
  required_aes = c("responses"),
  compute_panel = compute_panel_bar_delim,
  default_aes = ggplot2::aes(x = after_stat(cats), 
                             y = after_stat(num_responding))#,
  # setup_data = setup_data_function
  # compute_goup = compute_group_bar_delim
)

geom_bar_delim <- function(
  mapping = NULL,
  data = NULL,
  position = "identity",
  na.rm = FALSE,
  show.legend = NA,
  inherit.aes = TRUE, ...) {
  ggplot2::layer(
    stat = StatBardelim,  # proto object from Step 2
    geom = ggplot2::GeomCol,  # inherit other behavior
    data = data,
    mapping = mapping,
    position = position,
    show.legend = show.legend,
    inherit.aes = inherit.aes,
    params = list(na.rm = na.rm, ...)
  )
}
```

# try to use extension

```{r, eval  = F, error=T}
toy_survey %>%
  ggplot() +
  aes(responses = q1) +
  geom_bar_delim()
```


Backtrace:

  1. base (local) `<fn>`(x)

  2. ggplot2:::print.ggplot(x)

  4. ggplot2:::ggplot_build.ggplot(x)

  5. ggplot2:::by_layer(...)

 12. ggplot2 (local) f(l = layers[[i]], d = data[[i]])

     ...

 15. self$geom$setup_data(data, self$computed_geom_params)

 16. ggplot2 (local) setup_data(...)

 18. ggplot2::resolution(data$x, FALSE)

 20. base::Summary.factor(`<fct>`, na.rm = TRUE)

 21. base::stop(gettextf("%s not meaningful for factors", sQuote(.Generic)))

